---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, 
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message=FALSE, 
  warning=FALSE
)
```

Read the data containing daily case number of COVID-19 worldwide
```{r}
library(readr)
library(dplyr)
library(plotly)
library(tidyr)

covid19 <- read_csv("dataset/coronavirus.csv")
head(covid19)
```

This `coronavirus` dataset has the following fields:

* `date` - The date of the summary
* `province` - The province or state, when applicable
* `country` - The country or region name
* `lat` - Latitude point
* `long` - Longitude point
* `type` - the type of case (i.e., confirmed, death)
* `cases` - the number of daily cases (corresponding to the case type)


Plotting the total cases by type worldwide:

``` {r}
covid19 %>% 
  group_by(type, date) %>%
  summarise(total_cases = sum(cases)) %>%
  pivot_wider(names_from = type, values_from = total_cases) %>%
  arrange(date) %>%
  mutate(active = confirmed - death - recovered) %>%
  mutate(active_total = cumsum(active),
                recovered_total = cumsum(recovered),
                death_total = cumsum(death)) %>%
  plot_ly(x = ~ date,
                  y = ~ active_total,
                  name = 'Active', 
                  fillcolor = 'blue',
                  type = 'scatter',
                  mode = 'none', 
                  stackgroup = 'one') %>%
  add_trace(y = ~ death_total, 
             name = "Death",
             fillcolor = 'red') %>%
  add_trace(y = ~recovered_total, 
            name = 'Recovered', 
            fillcolor = 'green') %>%
  layout(title = "Distribution of Covid19 Cases Worldwide",
         legend = list(x = 0.1, y = 0.9),
         yaxis = list(title = "Number of Cases"),
         xaxis = list(title = "Source: Johns Hopkins University Center for Systems Science and Engineering"))
```

plot the confirmed cases distribution by counrty with treemap plot:

``` {r}
conf_df <- covid19 %>% 
  filter(type == "confirmed") %>%
  group_by(country) %>%
  summarise(total_cases = sum(cases)) %>%
  arrange(-total_cases) %>%
  mutate(parents = "Confirmed") %>%
  ungroup() 
  
  plot_ly(data = conf_df,
          type= "treemap",
          values = ~total_cases,
          labels= ~ country,
          parents=  ~parents,
          domain = list(column=0),
          name = "Confirmed",
          textinfo="label+value+percent parent")
```

Plot the daily cases from US

``` {r}
conf_US <- covid19 %>% 
  filter(type == "confirmed" & country == "US") %>%
  mutate(cum_cases = cumsum(cases)) %>%
  ungroup()
  
plot_ly(data = conf_US, x = ~date, 
        y = ~cases, type = "scatter", 
        mode = "line", name = "US_daily_cases", fillcolor = "blue")

```

Plot the cumulative_cases from US

``` {r}
plot_ly(data = conf_US, x = ~date, 
        y = ~cum_cases, type = "scatter", 
        mode = "line", name = "US_cumulative_cases", fillcolor = "blue") 

```

Plot cumulative cases from California in US

``` {r}
us_confirmed <- read_csv("dataset/time_series_covid19_confirmed_US.csv")


us_confirmed_cum_totay <- us_confirmed %>%
  select(-c(UID, iso2, iso3, code3, FIPS, Admin2, Country_Region, Lat, Long_, Combined_Key)) %>% 
  pivot_longer(-Province_State, names_to = "date", values_to = "cases") %>%
  group_by(Province_State, date) %>%
  summarise(total_cases = sum(cases)) %>%
  ungroup() %>%
  filter(date == max(date)) %>%
  arrange(-total_cases)

us_confirmed_state <- us_confirmed %>%
  select(-c(UID, iso2, iso3, code3, FIPS, Admin2, Country_Region, Lat, Long_, Combined_Key)) %>% 
  pivot_longer(-Province_State, names_to = "date", values_to = "cases") %>%
  group_by(Province_State, date) %>%
  summarise(total_cases = sum(cases)) %>%
  ungroup()


us_confirmed_CA <- us_confirmed_state %>%
  filter(Province_State == "California") 


us_confirmed_CA$date = format(as.Date(us_confirmed_CA$date, "%m/%d/%y"), "%m/%d")

us_confirmed_CA <- us_confirmed_CA %>%
  arrange(date)
plot_ly(data = us_confirmed_CA, x = ~date, y = ~total_cases, type = "scatter", mode = "line")


```

Plot daily cases from California in US

``` {r}
daily_mat <- matrix(0, dim(us_confirmed_CA)[1], 1)
colnames(daily_mat) <- 'CA_daily'

new <- cbind(us_confirmed_CA, daily_mat)

for (i in 2:dim(new)[1]){
  new[i,4] <- new[i,3] - new[i-1,3]
}

plot_ly(data = new, x = ~date, 
        y = ~CA_daily, type = "scatter", 
        mode = "line", name = "CA_daily", 
        fillcolor = "blue")
```